---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
pkgload::load_all(".")
```

# RcppMeCab

> This repo is a fork from [junhewk/RcppMeCab](https://github.com/junhewk/RcppMeCab).

This package, RcppMeCab, is an 'Rcpp' wrapper for the part-of-speech morphological analyzer MeCab. It supports native utf-8 encoding in C++ code and CJK (Chinese, Japanese, and Korean) MeCab library. This package fully utilizes the power Rcpp brings R computation to analyze texts faster.

## Usage

This package has `pos` and `posParallel` function.

```{r}
sentence <- "雨にも負けず、風にも負けず"

pos(sentence) # returns list, sentence will present on the names of the list
pos(sentence, join = FALSE) # for yielding morphemes only (tags will be given on the vector names)
pos(sentence, format = "data.frame") # the result will returned as a data frame format
posParallel(sentence) # parallelized version uses more memory, but much faster than the loop in single threading
```

+ sentence: a text for analyzing
+ join: If it gets TRUE, output form is (morpheme/tag). If it gets FALSE, output form is (morpheme) + tag in attribute.
+ format: The default is a list. If you set this as `"data.frame"`, the function will return the result in a data frame format.
+ sys_dic: a directory in which `dicrc` file is located, default value is `NULL` or you can set your default value using `options(mecabSysDic = "/path/to/your/system_dictionary")` 
+ user_dic: a user dictionary file compiled by `mecab_dict_index`, default value is also ""
